---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# biodivercity: Biodiversity indices for city landscapes

<!-- badges: start -->
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/biodivercity)](https://CRAN.R-project.org/package=biodivercity)
[![R-CMD-check](https://github.com/ecological-cities/biodivercity/workflows/R-CMD-check/badge.svg)](https://github.com/ecological-cities/biodivercity/actions)
<!-- badges: end -->

`biodivercity` is an R package for assessing the diversity of animal species in urban areas. Refer to [package website](https://ecological-cities.github.io/biodivercity/) for demonstrations of how the package may be used.  



## Installation

Install the development version of `biodivercity` from GitHub:

``` {r eval = FALSE}
# install.packages("devtools")
devtools::install_github("ecological-cities/biodivercity", ref = "main")
```


## Setup

Load the package:

```{r eval = FALSE}
library(biodivercity)
```

```{r include = FALSE, eval = TRUE}
devtools::load_all() # to knit manually w latest changes (not built/installed yet)
```

## Citation

To cite `home2park` or acknowledge its use, please cite the following:

*biodivercity: An R package to assess the diversity of animal species in urban areas. https://github.com/ecological-cities/biodivercity.*  


<br>

To get a BibTex entry, run `citation("biodivercity")`.

<br>
