---
title: "Process Animal Data"
subtitle: "Process animal survey data for model building"
author: "Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Process Animal Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Process survey data

Collected fauna survey data in the format detailed in the previous section can be easily subset with respect to `town`, `round` and/or `priority` with the `filter_observations()` function. For example, a specific set of surveys (i.e. butterfly surveys in Tampines from 2020-2021) can be filtered as follows:

```{r filter_obs example, eval = FALSE}
data(fauna)
data(surveys)

butterflies_tampines_round_2 <- 
  filter_observations(observations = fauna, survey_ref = surveys,
                      specify_taxon = "Lepidoptera",
                      specify_area = "TP",
                      specify_period = "2")
```

Checks are then performed to remove uncertain fauna identifications to prevent over-counting. 

```{r check taxon groups, eval = FALSE}

```
