---
title: "Animals: Survey Protocols"
subtitle: " "
author: "Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Animals: Survey Protocols}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Pre-survey preparations

### Sampling point generation
To generate random sampling points for the collection of new data, the research area is first randomly sampled at a specified point density. Land cover is stratified into two main classes of *natural vegetation* and *urban* land cover. These two land cover types are randomly sampled independently at a point density of one point per 50 hectares of land. 

```{r random point generation, eval = FALSE}
data(area)

points <- random_pt_gen()
```


### Data collection format

```{r sample data table, echo = FALSE, message = FALSE, warning = FALSE}

# show sample data in kable
library("dplyr")

# use either library (having installed package with latest changes) or load_all (package in development)
# library("biodivercity")
devtools::load_all() #knit manually with latest changes to circumvent library("biodivercity") (use in final version)

data(fauna)

knitr::kable(head(fauna), caption = "**Sample survey data format**") %>%
  kableExtra::kable_styling("striped", font_size = 10) %>% kableExtra::scroll_box(width = "100%")
```

```{r sample info table, echo = FALSE, message = FALSE}
# show sample info in kable
data(surveys)

knitr::kable(head(surveys), caption = "**Sample survey info format**") %>%
  kableExtra::kable_styling("striped", font_size = 10) %>% kableExtra::scroll_box(width = "100%")
```

